---
title: "Code for paper figures and stats"
output: 
  html_document:
    code_folding: 'hide'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Set up

```{r}
## Imports ----
library(tidyverse)
library(patchwork)
library(DT)
library(here)
source("./code/utils.R")

## Data ----
analytic_df <- read_csv("./data/analytic_data_wide.csv")
plotting_df <- process_to_plotting_df(analytic_df)
```

####  1. Percent over 70 (x) vs percent in poverty (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p70older > return_rounded_iqr(plotting_df$p70older)[2] &
            p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p70older)), sum(!is.na(p_poverty))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties.

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p70older)[2]` which includes `r temp_df %>% filter(p70older > return_rounded_iqr(plotting_df$p70older)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p70older)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p70older)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_poverty)[2]` which includes `r temp_df %>% filter(p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_poverty)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_poverty)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.

####  2. Percent over 70 (x) vs percent in group quarters (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p70older > return_rounded_iqr(plotting_df$p70older)[2] &
            p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p70older)), sum(!is.na(p_group_quarters))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties. 

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p70older)[2]` which includes `r temp_df %>% filter(p70older > return_rounded_iqr(plotting_df$p70older)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p70older)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p70older)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_group_quarters)[2]` which includes `r temp_df %>% filter(p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_group_quarters)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_group_quarters)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.

####  3. Percent in poverty (x) vs percent in group quarters (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2] &
            p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p_poverty)), sum(!is.na(p_group_quarters))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties. 

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_poverty)[2]` which includes `r temp_df %>% filter(p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_poverty)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_poverty)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_group_quarters)[2]` which includes `r temp_df %>% filter(p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_group_quarters)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_group_quarters)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.

####  4. Percent non-white (x) vs percent over 70 (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2] &
            p70older > return_rounded_iqr(plotting_df$p70older)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p_nonwhite)), sum(!is.na(p70older))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties. 

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_nonwhite)[2]` which includes `r temp_df %>% filter(p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_nonwhite)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_nonwhite)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p70older)[2]` which includes `r temp_df %>% filter(p70older > return_rounded_iqr(plotting_df$p70older)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p70older)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p70older)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.

####  5. Percent non-white (x) vs percent in poverty (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2] &
            p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p_nonwhite)), sum(!is.na(p_poverty))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties. 

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_nonwhite)[2]` which includes `r temp_df %>% filter(p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_nonwhite)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_nonwhite)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_poverty)[2]` which includes `r temp_df %>% filter(p_poverty > return_rounded_iqr(plotting_df$p_poverty)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_poverty)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_poverty)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.

####  6. Percent non-white (x) vs percent in group quarters (y)
```{r}
temp_df <- plotting_df %>%
    filter(!is.na(n_pop_2017)) %>%
    mutate(high_risk = ifelse(
        p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2] &
            p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2],
        1,
        0
    )) 

national_df <- temp_df %>%
    summarize(
        n_total_counties = sum(!is.na(n_pop_2017)),
        n_non_missing = min(sum(!is.na(p_nonwhite)), sum(!is.na(p_group_quarters))),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    )

national_df$pop_at_risk <- temp_df %>% 
  filter(high_risk == 1) %>% 
  pull(n_pop_2017) %>% 
  sum()

ranked_df <- temp_df %>%
    group_by(abbrev) %>%
    summarize(
        n_total_counties = n(),
        n_high_risk_counties = sum(high_risk, na.rm = TRUE),
        p_high_risk_counties = mean(high_risk, na.rm = TRUE)
    ) %>%
    left_join(
        temp_df %>%
            filter(high_risk == 1) %>%
            group_by(abbrev) %>%
            summarize(pop_at_risk = sum(n_pop_2017, na.rm = TRUE))
    ) %>% 
    mutate(n_rank = dense_rank(desc(n_high_risk_counties)), 
           p_rank = dense_rank(desc(p_high_risk_counties)),
           pop_rank = dense_rank(desc(pop_at_risk))) %>% 
    ungroup()
```

In the US, `r national_df$n_high_risk_counties` out of `r national_df$n_non_missing` counties (`r sprintf("%0.3f", round(national_df$p_high_risk_counties, 3))`) are at high risk for both. A total of `r sprintf("%s", round(national_df$pop_at_risk))` people live in these counties. 

For the x-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_nonwhite)[2]` which includes `r temp_df %>% filter(p_nonwhite > return_rounded_iqr(plotting_df$p_nonwhite)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_nonwhite)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_nonwhite)`.

For the y-variable, the 75th %ile value is `r return_rounded_iqr(plotting_df$p_group_quarters)[2]` which includes `r temp_df %>% filter(p_group_quarters > return_rounded_iqr(plotting_df$p_group_quarters)[2]) %>% NROW()` out of `r temp_df %>% filter(!is.na(p_group_quarters)) %>% NROW()` counties. The IQR is: `r return_rounded_iqr(plotting_df$p_group_quarters)`.

The top 5 states by number of counties are: `r ranked_df %>% filter(n_rank <= 5) %>% transmute(x = sprintf("%s (%i of %i; population at risk of %i)", abbrev, n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by proportion of counties are: `r ranked_df %>% filter(p_rank <= 5) %>% transmute(x = sprintf("%s (%0.3f; %i of %i counties; population at risk %i)", abbrev, round(p_high_risk_counties, 3), n_high_risk_counties, n_total_counties, pop_at_risk)) %>% pull(x)`.

The top 5 states by number of people affected are: `r ranked_df %>% filter(pop_rank <= 5) %>% transmute(x = sprintf("%s (%i people in %i of %i counties)", abbrev, pop_at_risk, n_high_risk_counties, n_total_counties)) %>% pull(x)`.
